---
title: "Male CNS Connectivity Analysis"
author: "Katherina Eichler, Modified FLorian Kämpf"
date: "2025-04-09"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This document performs an analysis of male CNS connectivity data. We
extract connectivity matrices, normalize the data using efficient sparse
matrix operations, and generate visualizations of input/output neuron
connectivity. The workflow includes: • Data extraction and preprocessing
• Construction and normalization of a sparse connectivity matrix •
Extraction and processing of neuron annotations • Layer propagation
computations • Aggregation and visualization via heatmaps and boxplots

## Load Required Libraries

```{r}
library(neuprintr)
library(arrow)
library(malecns)
library(bit64)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(Matrix)
library(cowplot)
library(coconat)  # for partner_summary2adjacency_matrix (if available)
```

# Data Extraction and Preprocessing

##  Extract Male CNS Annotations

```{r}
# Get male CNS annotations, filtering out unwanted statuses.
mdf <- mcns_dvid_annotations()
query.ids <- as.integer64(
  mdf$bodyid[!(mdf$status %in% c("Orphan", "PRT Orphan", "RT Orphan",
                                  "Unimportant", "Orphan hotknife",
                                  "Out of scope", "Glia", 
                                  "Orphan-artifact", NA))]
)
query_info <- mcns_neuprint_meta(query.ids)
query_info <- query_info[!(query_info$status %in% c("Orphan", NA)), ]
query.ids2 <- na.omit(as.character(query_info$bodyid))
rm(query.ids, query_info)
```

Load the Flat Connectome

```{r}
conn_feather <- arrow::open_dataset('/Users/fkampf/Downloads/connectome-weights-2023-11-15-f20f31-unlocked-minconf-0.5-primary-only.feather',
                                      format = 'feather')
conn_feather.inmem <- conn_feather %>%
  select(body_pre, body_post, weight) %>%
  collect()

#Create the Sparse Adjacency Matrix

all_ids <- union(unique(conn_feather.inmem$body_pre), unique(conn_feather.inmem$body_post))
np_adj <- coconat::partner_summary2adjacency_matrix(conn_feather.inmem, 
                                                    inputcol = 'body_pre', 
                                                    outputcol = 'body_post', 
                                                    sparse = TRUE,
                                                    inputids = all_ids, 
                                                    outputids = all_ids,
                                                    standardise_input = FALSE)
```

Matrix Scaling

Column Scaling Function

```{r}
#' Efficient column scaling for sparse matrices
#'
#' @description Normalizes a matrix's columns by dividing each column by its total sum.
#'   In connectivity data, columns represent downstream neurons. This function converts raw
#'   input weights into percentages.
#'
#' @param A A sparse (or dense) matrix.
#' @param na.rm Logical flag indicating whether to set non-finite scaling factors to 0.
#'
#' @return A matrix with each column scaled to sum to 1 (or zero if the original sum was zero).
#'
#' @examples
#' library(Matrix)
#' set.seed(42)
#' A <- Matrix(rbinom(100, 10, 0.05), nrow = 10)
#' colScale(A)
colScale <- function(A, na.rm = TRUE) {
  scalefac <- 1 / Matrix::colSums(A)
  if (na.rm) scalefac[!is.finite(scalefac)] <- 0
  B <- A %*% Matrix::Diagonal(x = scalefac)
  B
}
```

Compute the Input Percentage Matrix

```{r}
np_adj_per_in <- colScale(np_adj)
colnames(np_adj_per_in) <- colnames(np_adj)
```

Restrict to Valid Neuron Bodies

```{r}
np_adj <- np_adj[rownames(np_adj) %in% query.ids2, colnames(np_adj) %in% query.ids2]
np_adj_per_in <- np_adj_per_in[rownames(np_adj_per_in) %in% query.ids2, colnames(np_adj_per_in) %in% query.ids2]
rm(query.ids2, conn_feather.inmem, conn_feather, mdf)
```

Input/Output Neuron Annotations

Process Annotations

```{r}
all_info <- mcns_body_annotations()

cell_types <- c('CB1076', 'CB1078', 'CB3710', 'CB2521',
                'CB1542', 'CB1038', 'CB1427', 'CB2556', 'CB2380')
mba <- mcns_body_annotations()
mba.type.coalesced <- mba %>%
  mutate(type = coalesce(type, flywire_type, hemibrain_type, manc_type))
input_ids <- mba.type.coalesced %>% 
  filter(flywire_type %in% cell_types) %>%
  pull(bodyid) %>%
  unique()

input_info <- mba[mba$bodyid %in% input_ids, ]
input_info_sel <- input_info[, c("bodyid", "class", "group", "type", "instance", "soma_side", "root_side", "superclass")]
input_info_sel$soma_or_root_side <- ifelse(!is.na(input_info_sel$soma_side) & input_info_sel$soma_side != "",
                                            input_info_sel$soma_side,
                                            input_info_sel$root_side)

output_ids <- as.character(c(57956, 30261, 27091, 18554, 17337, 13495, 17918, 19136, 515347,
                             15035, 16719, 17089, 17589, 31586, 12442, 18276, 17098, 20597,
                             267996, 24537, 21349, 79094, 32788, 27072, 16559, 43592, 21586,
                             13655, 22744, 20803, 15324, 84181, 518589, 18830, 19894, 556999,
                             30669, 19818, 513129, 517178, 559181, 35694, 27507, 53355, 532582,
                             17867, 555488, 514909, 24072, 17618, 17851, 13790, 22346, 17434,
                             10930, 70011, 19948, 29992, 513515, 513651, 517653, 516954, 20392,
                             10765, 35993, 539247, 21507, 37651, 30273, 16805, 20117, 555751,
                             13562, 21033, 29043, 20068, 87497, 522357, 27330, 12889, 75407,
                             34082, 27368, 41366, 26543, 19600, 24874, 33318, 529875, 41518,
                             24060, 23968, 23558, 518253, 16541, 528986, 555811, 31905, 514077,
                             55707, 33009, 17841, 19109, 33503, 517337, 519518, 25542, 22093,
                             521453, 521449, 531383, 15876, 46660, 518038, 30391, 17702, 515990,
                             37522, 18861, 33435, 47325, 29462, 14492, 557035, 513652, 17487,
                             18017, 55136, 30558, 19200, 19236, 524297, 17470, 29483, 30455,
                             18990, 50692, 37977, 29143, 26989, 28094, 33519, 522419, 71385,
                             26746, 23319, 22270))
output_info <- all_info[all_info$bodyid %in% output_ids, ]
output_info_sel <- output_info[, c("bodyid", "class", "group", "type", "instance", "soma_side", "root_side", "superclass")]
output_info_sel$soma_or_root_side <- ifelse(!is.na(output_info_sel$soma_side) & output_info_sel$soma_side != "",
                                             output_info_sel$soma_side,
                                             output_info_sel$root_side)
```

Precalculate Layers and Compute Connectivity Scores

Precalculate Layers

```{r}
desired_layers <- 3
layer_vec <- paste0("v", seq(from = 0, to = desired_layers))
layer_list <- vector("list", length(layer_vec))
names(layer_list) <- layer_vec
layer_list[["v1"]] <- np_adj_per_in[rownames(np_adj_per_in) %in% input_info_sel$bodyid, ]
for (layer in 2:desired_layers) {
  layer_list[[layer + 1]] <- layer_list[[layer]] %*% np_adj_per_in
}
```

Compute Connectivity Scores

We compute the input/output connectivity scores and assign connectivity
details for each input neuron type.

```{r}
input_types <- unique(input_info_sel$type)
for (s in 1:length(input_types)) {
  assign(input_types[s], na.omit(unique(input_info$bodyid[grepl(input_types[s], input_info$type)])))
}
output_order <- sort(unique(output_info$type))

in_out_scores <- setNames(data.frame(matrix(ncol = 16, nrow = 0)),
                          c("bodyid", "conn_score", "ds_layer", "class", "group", "type", 
                            "instance", "soma_side", "rootSide", "superclass", 
                            "soma_or_root_side", "input", "in_side", "ipsi_contra"))
for (n in 1:length(input_types)) {
  startn_r <- as.character(na.omit(input_info_sel$bodyid[input_info_sel$type == input_types[n] & 
                                                            input_info_sel$soma_or_root_side == "R"]))
  startn_l <- as.character(na.omit(input_info_sel$bodyid[input_info_sel$type == input_types[n] & 
                                                            input_info_sel$soma_or_root_side == "L"]))
  if (length(startn_r) > 0 & length(startn_l) > 0) {
    input <- input_types[n]
    
    norm_list_l <- vector("list", length(layer_vec))
    names(norm_list_l) <- layer_vec
    norm_list_r <- vector("list", length(layer_vec))
    names(norm_list_r) <- layer_vec
    
    for (layer in 1:desired_layers) {
      if (length(startn_l) > 1) {
        temp_layer_conn_str <- Matrix::colSums(layer_list[[layer + 1]][na.omit(match(startn_l, rownames(layer_list[[layer + 1]]))), ])
      } else {
        temp_layer_conn_str <- layer_list[[layer + 1]][na.omit(match(startn_l, rownames(layer_list[[layer + 1]]))), ]
      }
      norm_list_l[[layer + 1]] <- setNames(as.vector(temp_layer_conn_str / mean(temp_layer_conn_str[temp_layer_conn_str > 0])),
                                           colnames(np_adj_per_in))
      if (length(startn_r) > 1) {
        temp_layer_conn_str <- Matrix::colSums(layer_list[[layer + 1]][na.omit(match(startn_r, rownames(layer_list[[layer + 1]]))), ])
      } else {
        temp_layer_conn_str <- layer_list[[layer + 1]][na.omit(match(startn_r, rownames(layer_list[[layer + 1]]))), ]
      }
      norm_list_r[[layer + 1]] <- setNames(as.vector(temp_layer_conn_str / mean(temp_layer_conn_str[temp_layer_conn_str > 0])),
                                           colnames(np_adj_per_in))
    }
    
    out_r_scores_df <- setNames(data.frame(matrix(ncol = 3, nrow = 0)),
                                c("bodyid", "conn_score", "ds_layer"))
    for (layer in 2:desired_layers) {
      out_r_scores <- norm_list_r[[layer]][names(norm_list_r[[layer]]) %in% output_ids]
      df <- data.frame(bodyid = as.double(names(out_r_scores)), conn_score = out_r_scores, ds_layer = layer - 1, row.names = NULL)
      out_r_scores_df <- rbind(out_r_scores_df, df)
    }
    out_r_scores_df %>%
      group_by(bodyid) %>%
      slice(which.max(conn_score)) -> out_r_scores_df_max
    
    out_r_scores_df_max_info <- left_join(out_r_scores_df_max, output_info_sel, by = "bodyid")
    out_r_scores_df_max_info$input <- input
    out_r_scores_df_max_info$in_side <- "R"
    
    out_l_scores_df <- setNames(data.frame(matrix(ncol = 3, nrow = 0)),
                                c("bodyid", "conn_score", "ds_layer"))
    for (layer in 2:desired_layers) {
      out_l_scores <- norm_list_l[[layer]][names(norm_list_l[[layer]]) %in% output_ids]
      df <- data.frame(bodyid = as.double(names(out_l_scores)), conn_score = out_l_scores, ds_layer = layer - 1, row.names = NULL)
      out_l_scores_df <- rbind(out_l_scores_df, df)
    }
    out_l_scores_df %>%
      group_by(bodyid) %>%
      slice(which.max(conn_score)) -> out_l_scores_df_max
    
    out_l_scores_df_max_info <- left_join(out_l_scores_df_max, output_info_sel, by = "bodyid")
    out_l_scores_df_max_info$input <- input
    out_l_scores_df_max_info$in_side <- "L"
    
    out_scores <- rbind(out_r_scores_df_max_info, out_l_scores_df_max_info)
    out_scores$ipsi_contra <- NA
    out_scores$ipsi_contra[out_scores$soma_or_root_side == out_scores$in_side] <- "ipsi"
    out_scores$ipsi_contra[out_scores$soma_or_root_side != out_scores$in_side] <- "contra"
    
    in_out_scores <- rbind(in_out_scores, out_scores)
  }
}
```

Aggregation and Visualization

Aggregate the Connectivity Scores

```{r}
in_out_scores %>%
  group_by(input, type, ipsi_contra) %>%
  summarise_at(vars("conn_score", "ds_layer"), mean) -> in_out_scores_mean

in_out_scores_mean %>% filter(ipsi_contra == "ipsi") -> in_out_scores_mean_ipsi
in_out_scores_mean_ipsi_m <- tidyr::pivot_wider(in_out_scores_mean_ipsi[, c("input", "conn_score", "type")],
                                                names_from = "input", values_from = "conn_score")
in_out_scores_mean_ipsi_m <- as.matrix(in_out_scores_mean_ipsi_m[, -1])
clust <- hclust(dist(t(in_out_scores_mean_ipsi_m)))
my_cols_fun <- colorRampPalette(c("#3E1F4B", "#8B008B", "#F2F2F2"))
```

Plot a Heatmap

```{r}
figure_heatmap <- ggplot(as_tibble(in_out_scores_mean), aes(x = factor(type, level = output_order), y = input)) +
  geom_point(aes(col = ds_layer, size = conn_score), shape = 15) +
  theme_minimal() +
  theme(legend.position = 'right', text = element_text(color = 'grey40')) +
  scale_size_area(max_size = 5) +
  scale_colour_gradientn(colours = my_cols_fun(desired_layers), limits = c(1, desired_layers)) +
  scale_y_discrete(limits = colnames(in_out_scores_mean_ipsi_m)[clust$order]) +
  guides(colour = guide_legend(override.aes = list(size = 10))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        text = element_text(size = 14, family = "Arial"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

figure_heatmap

Boxplots for Connectivity Scores

```{r}
for (nth in 1:length(input_types)) {
  in_out_scores_type <- in_out_scores[in_out_scores$input == input_types[nth] & in_out_scores$type %in% output_order, ]
  
  type_score <- ggplot(in_out_scores_type, aes(x = factor(type, level = output_order), y = conn_score)) +
    geom_boxplot(fill = "#8B008B") +
    ylab("Score") + xlab("Output Neurons") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
          panel.background = element_rect(fill = "white", colour = "black"),
          panel.grid = element_line(colour = "grey", linewidth = 0.2)) +
    ylim(0, ceiling(max(in_out_scores_type$conn_score))) +
    facet_grid(. ~ factor(ipsi_contra, levels = c("ipsi", "contra")), scales = "fixed")
  
  type_layer <- ggplot(in_out_scores_type, aes(x = factor(type, level = output_order), y = ds_layer)) +
    geom_boxplot(fill = "#8B008B") +
    ylab("Layer") + xlab("Output Neurons") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
          panel.background = element_rect(fill = "white", colour = "black"),
          panel.grid = element_line(colour = "grey", linewidth = 0.2)) +
    ylim(1, desired_layers) +
    facet_grid(. ~ factor(ipsi_contra, levels = c("ipsi", "contra")), scales = "fixed")
  
  legend <- cowplot::get_legend(
    type_score +
      guides(color = guide_legend(nrow = 1)) +
      theme(legend.position = "bottom")
  )
  title <- cowplot::ggdraw() +
    cowplot::draw_label(paste0(input_types[nth], " connectivity to leg MNs"), fontface = 'bold')
  
  figure_nwh <- cowplot::plot_grid(
    type_score + theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "none", plot.margin = margin(0.1,0.1,0,0.3, "cm")),
    type_layer + theme(legend.position = "none", strip.background = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0.1,0.1,0,0.3, "cm")),
    ncol = 1, common.legend = FALSE, rel_heights = c(1.3, 0.8), align = "v", vjust = 0, axis = "rlbt"
  )
  figure2_nwh <- cowplot::plot_grid(title, legend, nrow = 1, rel_widths = c(1,1))
  figure3_nwh <- cowplot::plot_grid(figure2_nwh, figure_nwh, ncol = 1, rel_heights = c(0.1,1))
  ggsave(paste0("output/", input_types[nth], "_combined_plot.pdf"),
         figure3_nwh, dev = cairo_pdf, width = 4715, height = 3295, units = "px")
}
```

Conclusion

This document demonstrates a full workflow for processing male CNS
connectivity data. We: • Extract and preprocess connectivity and
annotation data, • Normalize a large sparse connectivity matrix via
efficient scaling, • Compute layer-based connectivity scores, • And
visualize results using heatmaps and boxplots.

This report is intended to be reproducible and easily modifiable for
further analysis. Enjoy the analysis!
