

```{r}
fafbseg::flywire_connectome_data("syn")
library(malecns)
library(coconatfly)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(stringr)
library(igraph)
library(fafbseg)
library(tidygraph)
library(ggraph)
synapse_threshold=5
```

Just fetch some johnston organ neurons

```{r}

sensory.all<-cf_meta(cf_ids(malecns='superclass:sensory'))
sensory.johnston <- sensory.all%>%filter(grepl("JO-",type))
sensory.johnston.song <- sensory.all%>%filter(grepl("JO-A|JO-B",otype))
```

Look at postsynaptic partners of all sensory.johnston.song since there have to be aLN and aPN1

```{r}
sensory.johnston.song.output <-cf_partners(cf_ids(malecns=sensory.johnston.song$id),partners = 'output',threshold=5)
sensory.johnston.song.output.per.unique.cell <- sensory.johnston.song.output %>%
  group_by(partner)%>%
  summarize(total=sum(weight),type=first(type))
#12865 looks promising like aPN1 from 10.1016/j.cub.2014.03.048
```

Look at the inputs of aPN1 if there are more JO-B and if aLN(al) works with it

```{r}
input.CB2556 <- cf_partners(cf_ids(malecns='flywireType:CB2556'),partners = 'in',threshold=5) %>%
  left_join(cf_partners(cf_ids(malecns='flywireType:CB2556'),
                        threshold = synapse_threshold,
                        partners = 'in')%>%
              group_by(bodyid)%>%
              summarize(total.input.to.post=sum(weight))%>%
              as.data.frame(),
            by='bodyid')%>%
  left_join(cf_partners(cf_ids(malecns=cf_partners(cf_ids(malecns='flywireType:CB2556'),
                                                   threshold = synapse_threshold,
                                                   partners = 'in')%>%
                                 pull(pre_id)%>%
                                 unique()),
                        threshold = synapse_threshold,
                        partners = 'out')%>%
              group_by(pre_id)%>%
              summarize(total.output.by.pre=sum(weight))%>%
              as.data.frame(),
            by='pre_id')%>%
  mutate(fraction.of.output=weight/total.output.by.pre,
         fraction.of.input=weight/total.input.to.post)

input.CB2556.summary.across.type.post_id <- input.CB2556%>% 
  group_by(type,post_id) %>% 
  summarize(total=sum(weight),
            sum.fraction.output = sum(fraction.of.output),
            sum.fraction.input = sum(fraction.of.input),
            mean.fraction.output = mean(fraction.of.output),
            mean.fraction.input = mean(fraction.of.input))

input.CB2556.summary.across.type <- input.CB2556%>% 
  group_by(type) %>% 
  summarize(total=sum(weight),
            sum.fraction.output = sum(fraction.of.output),
            sum.fraction.input = sum(fraction.of.input),
            mean.fraction.output = mean(fraction.of.output),
            mean.fraction.input = mean(fraction.of.input))

input.CB2556.summary.across.unique.pre.cell <- input.CB2556%>% 
  group_by(pre_id) %>% 
  summarize(total=sum(weight),
            sum.fraction.output = sum(fraction.of.output),
            sum.fraction.input = sum(fraction.of.input),
            mean.fraction.output = mean(fraction.of.output),
            mean.fraction.input = mean(fraction.of.input),
            type=first(type))

input.CB2556.summary.across.unique.post.cell <- input.CB2556%>% 
  group_by(post_id) %>% 
  summarize(total=sum(weight),
            sum.fraction.output = sum(fraction.of.output),
            sum.fraction.input = sum(fraction.of.input),
            mean.fraction.output = mean(fraction.of.output),
            mean.fraction.input = mean(fraction.of.input),
            type=first(type))
```


Matching function

```{r}
prepare_data <- function(dataset.name,partner.df,post_or_pre){
  if (grepl(post_or_pre, 'post')) {post_or_pre='post_key'}
  else {post_or_pre='pre_key'}
  
  data <- cf_meta(partner.df%>%
                    filter(dataset==dataset.name)%>%
                    pull(post_or_pre)%>% #should be post right now
                    unique())%>%
    mutate(id=as.character(id))%>%
    rename_with(~ str_replace_all(.x, c(
      "flywireType" = "flywire",
      "flywire_type" = "flywire",
      "hemibrainType" = "hemibrain",
      "hemibrain_type" = "hemibrain",
      'malecns_type' = 'malecns',
      'malecns_type' = 'malecns',
      "type" = 'type2',
      'bodyid'='id'
    )))%>%
    select(any_of(c('id',
                    'type1',
                    'type2',
                    'flywire',
                    'hemibrain',
                    'malecns')))
  
  return(data)
}

make_edges <- function(data,sink_no=1){
  data %>% 
    pivot_longer(!id,
                 names_to = 'dataset',
                 values_to = "type")%>%
    select(id,type)%>%
    filter(!is.na(type))%>%
    mutate(sink=sink_no)}







match_cell_types <-function(partner.df,post_or_pre='post',flywire2mcns=T){
  #format data
  if (flywire2mcns){  
    data1 <- prepare_data('flywire', partner.df,post_or_pre)
    data2 <- prepare_data('malecns', partner.df,post_or_pre)}
  else {  
    data1 <- prepare_data('malecns', partner.df,post_or_pre)
    data2 <- prepare_data('flywire', partner.df,post_or_pre)}
  
  

  
  #concat data
  data.concat <- rbind(data1%>%
                         rename('type'='type2')%>%
                         select(id,type),
                       data2%>%
                         rename('type'='type2')%>%
                         select(id,type))
  
  #make edges
  edges1 <- make_edges(data1,'1')
  edges2 <- make_edges(data2,'2')
  edges<-rbind(edges1,edges2)
  
  #make nodes from edges
  nodes<-  edges %>% 
    pivot_longer(cols = c('id','type'), values_to = "name",names_to = 'type')%>%
    mutate(node_type=if_else(type=='type',
                             'cell_type',
                             str_c("sink_", sink)))%>%
    select(name, node_type)%>%
    distinct(name,.keep_all = T)
  
  #create graph
  graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)
  
  # Process graph distances
  graph.distances <- data.frame(distances(graph))
  colnames(graph.distances)<- nodes$name
  rownames(graph.distances)<- nodes$name
  
  graph.distances.sink2sink <- graph.distances[nodes$node_type=='sink_1',nodes$node_type=='sink_2']
  rownames(graph.distances.sink2sink)<- nodes$name[nodes$node_type=='sink_1']
  colnames(graph.distances.sink2sink)<- nodes$name[nodes$node_type=='sink_2']
  
  column.indices.shortest.path <- sapply(rownames(graph.distances.sink2sink), function(x) {
    temp <- graph.distances.sink2sink[x, , drop = FALSE]
    
    if(length(unique(temp)) == 1) {
      return(NA) 
    } else {
      min_col <- which(temp == min(temp), arr.ind = TRUE)[1,][2]
      
      if (temp[1, min_col] == Inf) {
        return(NA)
      }
      
      return(min_col)
    }
  })
  column.indices.shortest.path <- setNames(column.indices.shortest.path, rownames(graph.distances.sink2sink))
  
  matched.type.list<-list()
  
  
  for (x in rownames(graph.distances.sink2sink)) {
    
    
    
    if (is.na(column.indices.shortest.path[x])) {
      
      print('No matched type')
      matched.type = data.concat%>%filter(id==x)%>%pull(type)
      
    }
    else{
      node1 <- x
      node2 <- colnames(graph.distances.sink2sink)[column.indices.shortest.path[x]]
      
      # Get the shortest path once
      path <- shortest_paths(graph, from = node1, to = node2)
      
      matched.type <- V(graph)$name[path$vpath[[1]][length(path$vpath[[1]]) - 1]]
      if(length(matched.type)==0){readline()
        
      }
    }
    
    matched.type.list <- append(matched.type.list, list(matched.type))
    
  }
  matched.join.df <-  data.frame(id = rownames(graph.distances.sink2sink), match.type = as.character(matched.type.list))
  
  result<- rbind(data1%>%
                   left_join(matched.join.df,by='id')%>%
                   select(id,match.type),
                 data2%>%rename('match.type'='type2')%>%
                   select(id,match.type))
  return(result)
  
}



```

test
```{r}
cell_types <- c('CB1076', 'CB3710', 'CB2521', 'CB1425', 'CB1078', 'CB1038', 'CB1427', 'CB2556','CB2380','CB2556')  #ALL
id <-cf_ids(malecns=mba%>%
              filter(flywire_type %in% cell_types)%>%
              pull(bodyid))
cfp<-cf_partners(id,
                 partners='out',
                 threshold=5)%>%
  group_by(type)%>%
  summarize(total=sum(weight))
mcnst<-mcns_connection_table(id,
                             partners='out',
                             threshold=5)%>%
  group_by(type)%>%
  summarize(total=sum(weight))
```



Try make a postsynaptic difference plot for these neurons

```{r}
  #other potential aPN1 CB1076, CB3710, CB2521, CB1425, CB1078, CB1038
  #other potential aLN CB1427

cell_types <- c("CB3207", "CB2380", "CB2556")  #Conservative
cell_types <- c('CB1076', 'CB3710', 'CB2521', 'CB1425','CB1542', 'CB1078', 'CB1038', 'CB1427', 'CB2556','CB2380','CB2556')  #ALL
#cell_types <- c('CB2364', 'CB2763', 'CB2253', 'CB1614')  #asp k

#aln al 'CB2380', 'CB1427', 'CB3207')
#apn : 'CB1076', 'CB3710', 'CB2521', 'CB1425', 'CB1078', 'CB1038', 'CB2556','CB2556'

cell_types <- c('CB1076', 'CB3710', 'CB2521', 'CB1425', 'CB1078', 'CB1038', 'CB1427', 'CB2556','CB2380','CB2556')  #ALL

cf_list <- lapply(cell_types, function(ct) {
  cf_ids(malecns = paste0("flywireType:", ct),flywire = ct)
})

all.auditory.output <- cf_partners(do.call(c, cf_list), partners = 'out', threshold = synapse_threshold)%>%
  mutate(post_id=as.character(post_id))



#result.m2f <- match_cell_types(all.auditory.output, 'post',flywire2mcns=F)%>%rename('match.type.m2f'='match.type')
#all.auditory.output.match <- all.auditory.output %>%left_join(result.m2f,by=c('post_id'='id'))
#all.auditory.output$type <- all.auditory.output.match$match.type.m2f



first_non_zero <- function(x) {
  if (is.numeric(x)) {
    non_zero <- x[x != 0]
    if (length(non_zero) > 0) {
      return(non_zero[1])
    } else {
      return(0)
    }
  } else if (is.character(x)) {
    non_zero <- x[x != ""]  # for character columns, assuming empty strings are the "zero"
    if (length(non_zero) > 0) {
      return(non_zero[1])
    } else {
      return("")
    }
  } else {
    return(NA)  # Return NA for unsupported types
  }
}


# Extract unique post_keys from the dataset
auditory.downstream.unique.keys <- all.auditory.output %>%
  pull(post_key) %>%
  unique()

# Compute total input to each post
auditory.downstream.inputs <- cf_partners(auditory.downstream.unique.keys, 
                             threshold = synapse_threshold, 
                             partners = 'in') %>%
  mutate(post_id=as.character(post_id))%>%
  group_by(post_id) %>%
  summarize(total.input.to.post = sum(weight),
            .groups = 'drop')%>%
  left_join(cf_meta(auditory.downstream.unique.keys)%>%
              select(id,type,dataset),
            by=c('post_id'='id'))%>%
  ungroup()%>%
  group_by(type, dataset) %>%
  mutate(total.input.to.type = sum(total.input.to.post)) %>%
  ungroup()%>%
  rename('typo'='type')%>%
  select(post_id,
         total.input.to.type,
         total.input.to.post,
         typo)

# Get meta of downstream
auditory.downstream.flywire.meta <- cf_meta(all.auditory.output %>%
                          filter(dataset == 'flywire') %>%
                          pull(post_key) %>%
                          unique()) %>%
  select(id, malecns_type)

auditory.downstream.malecns.meta <- mcns_body_annotations() %>%
  select(bodyid, 
         dimorphism, 
         fru_dsx, 
         flywire_type, 
         synonyms) %>%
  mutate(bodyid = as.character(bodyid))

# Merge all information
all.auditory.type.summary <- all.auditory.output %>%
  mutate(post_id = as.character(post_id)) %>%
  left_join(auditory.downstream.inputs,
            by = "post_id") %>%
  mutate(post_id = as.character(post_id)) %>%
  left_join(auditory.downstream.flywire.meta, 
            by = c("post_id" = "id")) %>%
  left_join(auditory.downstream.malecns.meta, 
            by = c("post_id" = "bodyid")) %>%
  mutate(fraction.of.input.per.cell = weight / total.input.to.post,
         fraction.of.input.per.type = weight / total.input.to.type,
         type = coalesce(flywire_type, 
                         malecns_type, 
                         type)) %>%
  group_by(type, dataset) %>%
  summarize(
    total = sum(weight),
    mean.fraction.of.input.per.cell = mean(fraction.of.input.per.cell, 
                                  na.rm = TRUE),
    sum.fraction.of.input.per.cell = sum(fraction.of.input.per.cell, 
                                na.rm = TRUE),
    mean.fraction.of.input.per.type = mean(fraction.of.input.per.type, 
                                  na.rm = TRUE),
    sum.fraction.of.input.per.type = sum(fraction.of.input.per.type, 
                                na.rm = TRUE),
    n=n(),
    dimorphism = first(dimorphism),
    fru_dsx = first(fru_dsx),
    synonyms = first(synonyms),
    .groups = 'drop') %>%
  complete(type, 
           dataset, 
           fill = list(total = 0)) %>%
  arrange(desc(total)) %>%
  pivot_wider(
    id_cols = c(type, 
                dimorphism, 
                fru_dsx, 
                synonyms),
    names_from = dataset,
    values_from = c(total, 
                    mean.fraction.of.input.per.cell, 
                    sum.fraction.of.input.per.cell,
                    mean.fraction.of.input.per.type, 
                    sum.fraction.of.input.per.type,
                    n)
  ) %>%
  mutate(
    across(where(is.numeric), 
           ~ replace_na(.x,
                        0)), # Replace NA with 0 for numeric columns
    across(where(is.character), 
           ~ replace_na(.x,
                        "Unknown")) # Replace NA with "Unknown" for character columns
  ) %>%
  group_by(type) %>%
  summarize(across(everything(), first_non_zero), 
            .groups = 'drop') %>%
  mutate(m_f_ratio = total_malecns / total_flywire)



```

Scatter plot of inputs of male vs female auditory neurons

```{r}


p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = total_malecns, 
                                         y = total_flywire,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Absolute output to types by auditory aLNs and aPN1s",
       x = "total_malecns",
       y = "total_flywire",
       color = "Dataset") +
  geom_abline()+
  theme_minimal()

ggplotly(p, tooltip = "text")


p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = total_malecns/n_malecns, 
                                         y = total_flywire/n_flywire,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Absolute output to types by auditory aLNs and aPN1s",
       x = "total_malecns normed",
       y = "total_flywire normed",
       color = "Dataset") +
  geom_abline()+
  theme_minimal()

ggplotly(p, tooltip = "text")




p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = total_malecns-total_flywire, 
                                         y = m_f_ratio,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Absolute output to types by auditory aLNs and aPN1s",
       x = "delta (mcns-fafb)",
       y = "ratio (mcns/fafb)",
       color = "Dataset") +
  theme_minimal()

ggplotly(p, tooltip = "text")

flytable<-flytable_query("select * from info")

p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = (total_malecns/n_malecns)-(total_flywire/n_flywire), 
                                         y = m_f_ratio,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Absolute output to types by auditory aLNs and aPN1s",
       x = "delta normed by n of type (mcns-fafb)",
       y = "ratio (mcns/fafb)",
       color = "Dataset") +
  theme_minimal()

ggplotly(p, tooltip = "text")


p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = mean.fraction.of.input.per.type_malecns,
                                         y = mean.fraction.of.input.per.type_flywire,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Mean fraction output to types by auditory aLNs and aPN1s",
       x = "malecns",
       y = "flywire",
       color = "Dataset") +
  geom_abline()+
  theme_minimal()
ggplotly(p, tooltip = "text")


p<-ggplot(all.auditory.type.summary%>%mutate(dimorphism=if_else(dimorphism==0,NA,dimorphism)), aes(x = sum.fraction.of.input.per.type_malecns,
                                         y = sum.fraction.of.input.per.type_flywire,
                                         text=type,
                                         color = fru_dsx)) +
  geom_point(size = 1, alpha = 0.7) +
  theme_minimal() +
  labs(title = "sum fraction output to types by auditory aLNs and aPN1s",
       x = "malecns",
       y = "flywire",
       color = "Dataset") +
  geom_abline()+
  theme_minimal()
ggplotly(p, tooltip = "text")

```




See where the input to vpoIN in female comes from


```{r}


library(fafbseg)
mba<-mcns_body_annotations()
flytable_reduced<-flytable_query("select * from info")%>%select(root_id,cell_class,cell_type,hemibrain_type,malecns_type)%>%rename('cell_class_post'='cell_class','hemibrain_type_post'='hemibrain_type','malecns_type_post'='malecns_type')
input.vpoIN.level1<-cf_partners(c(cf_ids(flywire='AVLP008'),
              cf_ids(flywire='CB1385'),
              cf_ids(flywire='LHAV4c2'),
              cf_ids(flywire='LHAV4j1')),
            partners='in',threshold=5)%>%
            mutate(post_id=as.character(post_id))%>%
            left_join(flytable_reduced,by=c('post_id'='root_id'))%>%
            group_by(type)%>%
            summarize(total=sum(weight),
            post_cell_class=first(cell_class_post),
            post_cell_type=first(cell_type),
            post_malecns_type=first(malecns_type_post),
            post_hemibrain_type=first(hemibrain_type_post))%>%
            arrange(desc(total))
level.types<- (input.vpoIN.level1%>%pull(type))[1:10]%>%as.list
cf_list_flywire <- lapply(level.types, function(ct) {
  cf_ids(flywire = ct)
})

input.vpoIN.level2<-cf_partners(do.call(c, cf_list_flywire),
                                partners='in',
                                threshold=5)%>%
            mutate(post_id=as.character(post_id))%>%
            left_join(flytable_reduced,by=c('post_id'='root_id'))%>%
            group_by(type)%>%
            summarize(total=sum(weight),
            post_cell_class=first(cell_class_post),
            post_cell_type=first(cell_type),
            post_malecns_type=first(malecns_type_post),
            post_hemibrain_type=first(hemibrain_type_post))%>%
            arrange(desc(total))
            
level.types<- (input.vpoIN.level2%>%pull(type))[1:10]%>%as.list
cf_list_flywire <- lapply(level.types, function(ct) {
  cf_ids(flywire = ct)
})
input.vpoIN.level3<-cf_partners(do.call(c, cf_list_flywire),
                                partners='in',
                                threshold=5)%>%
                                group_by(type)%>%
            summarize(total=sum(weight))%>%
            arrange(desc(total))

```

Make a connectitiy matrix for the circuit flywire

```{r}


library(pheatmap)
library(reshape2)

cell_types.level1 <- c('CB1076', 'CB3710', 'CB2521', 'CB1425', 'CB1078', 'CB1038', 'CB1427', 'CB2556','CB2380','CB2556')

#level 1 fw
cf_list.level1.fw <- lapply(cell_types, function(ct) {
  cf_ids(flywire = ct)
})
level1.fw.cm <- cf_partner_summary(do.call(c, cf_list.level1.fw), 
                                          partners = 'out', 
                                          threshold = synapse_threshold,
                                          normalise = T, 
                                          aggregate.query = T)%>%
  arrange(desc(weight))%>%
  filter(weight>0.025)


#level 2 fw
cell_types.level2.fw <- c(cell_types.level1,
                          (level1.fw.cm%>%filter(!type.post %in% cell_types.level1))[1:10,]%>%pull(type.post))
cf_list.level2.fw <- lapply(cell_types.level2.fw, function(ct) {
  cf_ids(flywire = ct)
})

level2.fw.cm <- cf_partner_summary(do.call(c, cf_list.level2.fw), 
                                          partners = 'out', 
                                          threshold = synapse_threshold,
                                          normalise = T, 
                                          aggregate.query = T)%>%
  arrange(desc(weight))%>%
  filter(weight>0.025)

#effective calc
connected.init.lv2 <- level2.fw.cm%>%filter(type.pre %in% cell_types.level1) %>% pull(type.post)

#level 3 fw
cell_types.level3.fw <- cell_types.level2.fw <- c(cell_types.level2.fw,
                                                  (level2.fw.cm%>%filter(!type.post %in% cell_types.level2.fw))[1:10,]%>%pull(type.post))

cf_list.level3.fw <- lapply(cell_types.level3.fw, function(ct) {
  cf_ids(flywire = ct)
})

level3.fw.cm <- cf_partner_summary(do.call(c, cf_list.level3.fw), 
                                          partners = 'out', 
                                          threshold = synapse_threshold,
                                          normalise = T, 
                                          aggregate.query = T)%>%
  arrange(desc(weight))
 
level3.fw.cm <- level3.fw.cm %>% filter(type.post %in% (level3.fw.cm%>%filter(weight>0.01)%>%pull(type.post)))
conn_matrix <- dcast(level3.fw.cm, type.pre ~ type.post, value.var = "weight", fill = 0)

rownames(conn_matrix) <- conn_matrix$type.pre
conn_matrix <- conn_matrix[, -1]  # Remove the type.pre column as it's now row names
library()
pheatmap(conn_matrix, cluster_rows = T, cluster_cols = T, display_numbers = F,fontsize_col = 3,angle_col=90)




```
connectivity matrix malecns

```{r}


library(pheatmap)
library(reshape2)

cell_types.level1 <- c('CB1076', 'CB3710', 'CB2521', 'CB1425', 'CB1078', 'CB1038', 'CB1427', 'CB2556','CB2380','CB2556')

#level 1 fw
cf_list.level1.mcns <- lapply(cell_types, function(ct) {
  cf_ids(malecns = paste0("flywireType:", ct))
})


level1.fw.cm <- cf_partner_summary(do.call(c, cf_list.level1.mcns), 
                                          partners = 'out', 
                                          threshold = synapse_threshold,
                                          normalise = T, 
                                          aggregate.query = T)%>%
  arrange(desc(weight))%>%
  filter(weight>0.01)






```
